\documentclass[t]{beamer}

\subtitle{Section 5.1 Irreducibles and Unique Factorization}

\input{../../_tools/setup}

\begin{document} 
	\startdoc
	
	\topics{
		\item Algebraic structure of quotients of polynomial rings
	}
	{
		\item Factorization
		\item Divisibility in Rings
		\item Irreducible Ring Elements
	}

\slide{
	\begin{block}{\textbf{Motivation.}}
		Consider $R=\Z$, factor $a=70$ into prime elements.\vskip 2in\mbox{}
	\end{block}
	\begin{block}{\textbf{Goal.}}
		What do primes mean in other rings what does it mean to factor elements?
	\end{block}
}
\slide{
	\begin{defn}
		Let $R$ be an integral domain (with unity) and $a,b\in R$.  We say $a$ \emph{divides} $b$ and write $a|b$ if $ac=b$ for some $c\in R$.
	\end{defn}
	\begin{block}{\textbf{Some Properties.}}
		Let $R$ be an integral domain and $a,b,c\in R$.
		\enumarabic{\setlength{\itemsep}{1em}
			\item $a|a$
			\item if $a|b$ and $b|c$, then $a|c$
			\item if $a|b$ and $a|c$, then $a|(br+cs)$ for all $r,s\in R$
		}
	\end{block}
}
\slide{
	\begin{nb}
		In $R=\Z$, if $a|b$ and $b|a$, then \blank{1in}.
	\end{nb}
	\begin{thm}{5.1.1}
		If $R$ is an integral domain the following are equivalent for $a,b\in R$:\enumarabic{
			\item $a|b$ and $b|a$
			\item $a=ub$ for some unit $u\in R^*$
			\item $(a)=(b)$
		}
	\end{thm}
}
\slide{\vskip -.25in
	\begin{proof}
		Let $R$ be an integral domain and $a,b\in R$.
		
		$(1)\Rightarrow(2)$ Assume $a|b$ and $b|a$.\vskip 1in
		$(2)\Rightarrow(3)$ Assume $a=ub$ for some unit $u\in R^*$\vskip 1in
		$(3)\Rightarrow(1)$ Assume $(a)=(b)$.\vskip 1in
	\end{proof}
}
\slide{
	\begin{defn}
		We call $a,b\in R$ an integral domain \emph{associates} if $a|b$ and $b|a$.  We may write $a\sim b$.
	\end{defn}
	\begin{exa}
		Show $\sqrt{3}\sim(3+2\sqrt{3})$ in $R=\Z[\sqrt{3}]$.
		
		Book has one method, let's attempt another. Consider
		
		$$\frac{\sqrt{3}}{3+2\sqrt{3}}=\hspace*{2in}$$
		\vskip 2in
	\end{exa}
}
\slide{
	\begin{defn}
		We call an element $p$ in the integral domain $R$ \emph{irreducible} if
		\enumarabic{\item $p\neq0$ and $p$ is not a unit\item If $p=ab$ in $R$, then $a$ or $b$ is a unit in $R$.}
		Anything that is not irreducible is called \emph{reducible}.
	\end{defn}
}


\slide{
\begin{exa}
	Let's talk $\Z[\sqrt{d}]$ where $d\in\Z$ is \emph{squarefree}, that is $p^2\nmid d$ for all primes $p$.
	
	\bulletize{\setlength{\itemsep}{3em}
		\item Norm:
		\item Norm is Multiplicative:
		\item Units:
	}
\end{exa}
}
\slide{
\begin{exercise}
	Show that $4+\sqrt{5}$ is irreducible in $\Z[\sqrt{5}]$.
	\vskip 1in
	Do the same for $1+\sqrt{5}$.
	\vskip 2in\mbox{}
\end{exercise}
}

\slide{
	\begin{thm}{5.1.2}
		If $R$ is an integral domain, the following conditions are equivalent for a nonzero nonunit $p$ in $R$.
		\enumarabic{
			\item $p$ is irreducible.
			\item If $d|p$, then $d\sim 1$ or $d\sim p$.
			\item If $p\sim ab$ in $R$, then $\sim a$ or $p\sim b$.
			\item If $p=ab$ in $R$, then $p\sim a$ or $p\sim b$.
		}
	\end{thm}
	\begin{proof}
		\vskip 2in
	\end{proof}
}
\slide{
	\begin{statementblock}{Corollary}
		If $p\sim q$ in an integral domain $R$, then
		\begin{center} $p$ is irreducible if and only if $q$ is irreducible\end{center}
	\end{statementblock}
}
\slide{
	\begin{defn}
		We call an element $p$ of an integral domain $R$ \emph{prime} if
		\enumarabic{
			\item $p$ is nonzero and not a unit
			\item If $p|ab$ in $R$, then $p|a$ or $p|b$.
		}
	\end{defn}
	\begin{thm}{5.1.5}
		If $R$ is an integral domain and $p$ is prime, then $p$ is irreducible.
	\end{thm}
}


\slide{
	\begin{thm}{5.1.5}
		If $R$ is an integral domain and $p$ is prime, then $p$ is irreducible.
	\end{thm}
	\begin{block}{\textbf{Warning.}}
		The converse does not hold!  For example $p=1+\sqrt{-5}$ in $R=\Z[\sqrt{-5}]$.
		\bulletize{
			\item $p$ is irreducible:\vskip .75in
			\item $p$ is not prime:\vskip .75in\mbox{}
		}
	\end{block}
}
\slide{
	\begin{defn}
		An integral domain $R$ is called a \emph{unique factorization domain} if is satisfies both
		\enumarabic{
			\item Every nonzero nonunit in $R$ is a product of irreducibles.
			\item If $p_1p_2\cdots p_r\sim q_1q_2\cdots q_s$ where the $p_i$ and $q_j$ are irreducibles, then $r=s$ and (after possible relabeling) $p_i\sim q_i$ for each $i$.
		}
	\end{defn}
	\begin{statementblock}{Lemma 5.1.3}
		If $R$ is a UFD, then every irreducible in $R$ is prime.
	\end{statementblock}
}
\slide{
	\begin{exa}
		\bulletize{
			\item $\Z[i]$ is a UFD
			\item $\Z\left[\frac{-1+i\sqrt{3}}{2}\right]=\Z[e^{2\pi i/3}]$ is a UFD, we call this set the \emph{Eistenstein integers}
			\item $\Z[\sqrt{-5}]$ is not a UFD
			\item $d$ positive squarefree integer, $\Z[\sqrt{-d}]$ is a UFD  if and only if $d=1,2,3,7,11,19,43,67,163$, the so-called ``Heegner Numbers''}
	\end{exa}
}

\slide{
	\begin{thm}{5.1.10}
		If $R$ is a UFD, then so is $R[x]$.
	\end{thm}
	\begin{statementblock}{Corollary}
		If $R$ is a UFD, then so is $R[x_1,x_2,\dots,x_n]$.
	\end{statementblock}
}

\slide{
	\begin{exercise}
		If $R[x]$ is a UFD does that mean $R$ is too?\vskip 3in\mbox{}
	\end{exercise}
}
\slide{
	\begin{defn}
		Let $R$ be an integral domain and $s_1,s_2,\dots,s_n\in R$. A \emph{greatest common divisor} of $s_1,s_2,\dots,s_n$, denoted
		\[\gcd(s_1,s_2,\dots,s_n),\]
		is an element $d\in R$ such that
		\enumarabic{
			\item $d\mid s_i$ for all $i$
			\item If $r\in R$ satisfies $r|s_i$ for all $i$, then $r|d$ too.
		}
	\end{defn}
	\begin{nb}
		This element $d$ is unique only up to associates.  
	\end{nb}
}
\slide{
	\begin{defn}
		Let $R$ be an integral domain and $s_1,s_2,\dots,s_n\in R$. A \emph{least common multiple} of $s_1,s_2,\dots,s_n$, denoted
		\[\operatorname{lcm}(s_1,s_2,\dots,s_n),\]
		is an element $\ell\in R$ such that
		\enumarabic{
			\item $s_i|\ell$ for all $i$
			\item If $r\in R$ satisfies $s_i|r$ for all $i$, then $\ell|r$ too.
		}
	\end{defn}
	\begin{nb}
		This element $\ell$ is unique only up to associates.  
	\end{nb}
}

\end{document}

		