\documentclass[t]{beamer}

\subtitle{Section 0.2: Sets}

\input{../../_tools/setup}

\begin{document} 
\startdoc

\topics{
	% Last Time
	\item Course Introduction
	\item Direct Proofs
	\item Cases
}
{
	% This time
	\item Contradiction
	\item Counterexamples
	\item Contrapositive/Converse
	\item Sets and Operations
	\item Cardinality
	\item Power Sets
	\item Cartesian Products
}
\slide{
\begin{block}{Method of Proof by Contradiction.}
	To prove ``if $P$ then $Q$'' using contradiction, assume $P$ is true and $Q$ is false then show that something is terribly wrong.
	\begin{exa}
		Prove that there are no integers $x$ and $y$ satisfying $24x+12y=1$.
		%\enumarabic{\item If $x$ and $y$ are integers then $24x+12y=1$.}
	\end{exa}
\end{block}
}
\slide{
\begin{block}{Proof.}
	Let $x$ and $y$ be integers.  Assume, toward contradiction, that $24x+12y=1$.
	%\pause
	%Divide both sides of the equation $24x+12y=1$ by $12$ to get
	%	\[2x+y=\frac{1}{12}.\]\pause
	%Notice that since $x$ and $y$ are integers, so is $2x+y$.  However $\frac{1}{12}$ is not an integer, contradicting that fact that $2x+y=\frac{1}{12}$.
	
	%Therefore, the assumption that $24x+12y=1$ cannot be true.
\end{block}	
}
\slide{
	\begin{exercise}
		Using the last proof as inspiration, what do you think must be true about $m$ and $n$ if 	\[mx+ny=1\]
		is satisfied for integers $x$ and $y$?
		
		Thank about what property(ies) $m=24$ and $n=12$ had that caused a problem.  Why does $m=4$ and $n=9$ have an integer solution with $x=-2$ and $y=1$?
	\end{exercise}
}
\slide{
	\begin{block}{Counterexamples.}
		Sometimes we can prove a universal statement is false by finding a simple counterexample.
		\vskip 1cm
		\begin{exa}
			Prove the result or give a counterexample.
			\begin{quote}
				If $n^2=1$ then $n=1$.
			\end{quote}
			\pause
			\begin{block}{Solution.}
				This statement is false because if $n=-1$ then $n^2=1$ while $n\neq 1$.
			\end{block}
		\end{exa}
	\end{block}
}
\slide{
	\begin{defn}
		The \emph{contrapositive} of the statement $P\Rightarrow Q$ is $\sim Q\Rightarrow \sim P$.
		
		The \emph{converse} of the statement $P\Rightarrow Q$ is $Q\Rightarrow P$.
	\end{defn}
	\vskip 1cm
	\begin{nb}
		The contrapositive of a statement is logically equivalent to it - and is sometimes easier to prove!
		
		The converse of a statement is NOT logically equivalent to it.  
	\end{nb}
}
\slide{
	\begin{exercise}
		\enumarabic{
			\item If $n$ is an even integer then $n^2$ is a multiple of 4.
			\enumalph{
				\item State the converse and the contrapositive. 
				\item Which of the converse and contrapositive are true?
				\item If either is false, give a counterexample.
			}
			\item If $m+n=25$, where $n$ and $m$ are integers, then\\exactly one of $n$ and $m$ is greater than $12$.
			\enumalph{
				\item State the converse and the contrapositive. 
				\item Which of the converse and contrapositive are true?
				\item If either is false, give a counterexample.
			}
			\item For all integers $n$, if $n\geq 2$, then $n^3\geq 2^n$.
			\enumalph{
				\item State the converse and the contrapositive. 
				\item Which of the converse and contrapositive are true?
				\item If either is false, give a counterexample.
			}
		}
	\end{exercise}
}
\slide{
	\begin{defn}
		A \emph{set} is a collection of objects called \emph{elements}.
	\end{defn}
	\begin{block}{\textbf{Notation.}}
		If $S$ is a set with an element $a$, then we write $a{\color{blue}\in} S$.
	\end{block}
	\begin{exa}
		\bulletize{
			\item $\Z$
			\item $\R$
			\item $\C$
			\item $\Q$
			\item $\N$
			\item $\Z^+$
		}
	\end{exa}
}
\slide{
	\begin{exa}
		Sets can be represented in many ways:
		\bulletize{
			\item Listing all elements: $\{1,4,7,322\}$
			\item Giving a formula: $\{3a-2\in\Z\ |\ a\in\Z\}$
			\item Giving a criteria: $\{a\in\R\ :\ 2^a>5\}$
		}
	\end{exa}
	\begin{exercise}
		Determine the elements of the following sets:
		\enumalph{\setlength{\itemsep}{1em}
			\item $\{n\in \Z\ :\ |n-2|\leq 1\}$
			\item $\{10k\ :\ k\in \Z\}$
			\item $\{m+1\ |\ m\in \N \text{ and }m\text{ is even}\}$
		}
	\end{exercise}
}
\slide{
	\begin{defn}
		If $A$ has $n$-elements then we say the \emph{cardinality} of $A$ is $n$ and we write $|A|=n$. Such sets are called \emph{finite} sets.  
		Sets with an infinite number of elements are \emph{infinite} sets.
	\end{defn}
	\begin{exa}
		$|\emptyset|=0$
	\end{exa}
	\begin{exercise}
		Compute $|A|$ and $|B|$ given
		\[A=\{1,\{2\}\}\text{ and }B=\{\emptyset,\{1\},\{2\},\{1,2\}\}.\]
	\end{exercise}
}
\slide{
	\begin{defn}
		For sets $A$ and $B$ we say that \emph{$A$ is a subset of $B$} if every element of $A$ is also an element of $B$. In mathematical notation we would write
		\[\forall a\in A,\ a\in B.\]
	\end{defn}
	\begin{block}{\textbf{Notation.}}
		If $A$ is a subset of $B$ we write $A\ {\color{blue}\subseteq}\ B$.
	\end{block}
}
\slide{
	\begin{exercise}
		Let $A=\{x\in\R\ |\ x^2-1<1\}$ and $B=\{x\in\R\ |\ x<3\}$. Show that $A\subseteq B$.
	\end{exercise}
}
\slide{
	\begin{defn}
		The \emph{power set} of a set $A$ is the set $P(A)$ consisting of all subsets of $A$.
	\end{defn}
	\begin{exa}
		$B=\{\emptyset,\{1\},\{2\},\{1,2\}\}=P(\{1,2\})$
	\end{exa}
	\begin{exercise}
		If $A=\{4,5,6\}$, what is $P(A)$ and $|P(A)|$?
	\end{exercise}
}
\slidec{
	\begin{statementblock}{Principle of Set Equality}
		If $A$ and $B$ are sets then
		\begin{center}
			$A=B$\quad if and only if \quad $A\subseteq B$ and $B\subseteq A$
		\end{center}
	\end{statementblock}
}
\slide{
	\begin{defn}
		The \emph{intersection} of the sets $A$ and $B$, denotes $A\cap B$ is the set of all elements that appear in both $A$ and in $B$.
	\end{defn}
	\begin{defn}
		The \emph{union} of the sets $A$ and $B$, denotes $A\cup B$ is the set of all elements that appear in at least one of $A$ or $B$.
	\end{defn}
	\begin{exercise}
		Use the Principle of Set Equality to show that for sets $A$, $B$, and $C$, 
			\[A\cap (B\cup C) = (A\cap B)\cup (A\cap C).\]
	\end{exercise}
}
\slide{
	\begin{defn}
		The \emph{Cartesian Product} of the sets $A$ and $B$ is the set
		\[A\times B:=\{(a,b):a\in A\text{ and }b\in B\}.\]
		Note that the elements, $(a,b)$, are ordered pairs.
	\end{defn}
	\begin{exercise}
		Compute $A\times B$ if $A=\{3,4\}$ and $B=\{a,b,c\}$.
	\end{exercise}
}
\slide{
	\begin{exercise}
		\enumalph{
			\item Compute $A\times A$ if $A=\{2,4,6\}$.\vskip 1in
			
			\item Consider $A=[-2,6)$ and $B=(5,9)$ as intervals in $\R$.  What is $A\times B$?}
	\end{exercise}
}
\end{document}

