\documentclass[12pt]{article}

\newcommand{\secname}{Section 2.1: Binary Operations}

\input{../../_tools/handout_setup}
\begin{document}
	\startdoc

	\topics{
		\item Disjoint Cycles
		\item Transpositions
		\item Even vs Odd Permutations
	}{
		\item Binary Operations
		\item Monoids
	}

\slide{
	\begin{defn}
		A \emph{binary operation}, $*$ on a set $S$ is a function that associates to each ordered pair $(a,b)\in S\times S$ an element of $S$ which we call $a*b$.
	\end{defn}
	\begin{nb}
		Since we know that $a*b\in S$ for all $a,b\in S$, we say that the binary operation is \emph{closed under $*$}.
	\end{nb}
%	\begin{exa}\begin{itemize}[label=$\bullet$]
%			\item $+,-,\cdot,\div$ are all binary operations on $\R$
%			\item $+,-,\cdot$ are all binary operations on $\Z$ and on $\Z_n$
%			\item $\circ$ is a binary operation on $S_n$
%		\end{itemize}
%	\end{exa}
	\begin{nb}
		 Sometimes we write $(S,*)$ to mean $*$ is a binary operation on $S$.
	\end{nb}
}

\slide{
	\begin{defn}
		A binary operation $*$ on $S$ is \emph{associative} if
		\[a*(b*c)=(a*b)*c,\]
		for all $a,b,c\in S$.
	\end{defn}
}
\slide{
	\begin{defn}
		A binary operation $*$ on $S$ is \emph{commutative} if
		\[a*b=b*a,\]
		for all $a,b\in S$.
	\end{defn}
}
\slide{
	\begin{defn}
		An element $e\in S$ is called an \emph{identity} (or \emph{unity}) for the binary operation $*$ if
			\[a*e=e*a=a,\]
		for all $a\in S$.
	\end{defn}
}
\slide{
	\begin{statementblock}{Theorem 2.1.1}
		If a binary operation $*$ on a set $S$ has an identity, then it is unique.
	\end{statementblock}
}

\slide{
	\begin{defn}
		A set $S$ along with a binary operation $*$ is called an \emph{monoid} if $*$ is associative and has an identity.

		\vskip 1em

		If $(S,*)$ is also commutative, then we say $S$ is a \emph{commutative monoid}.
	\end{defn}
	\begin{exa}
		\bulletize{
			\item $(\Z,\cdot)$
			\item $(S_n,\circ)$
		}
	\end{exa}
}
\begin{exercise}
	Let $*$ be the binary operation on $\mathbb{N}$ given by \[a*b:=a^b.\]
	\enumalph{
		\item Associative?
		\vskip 2em
		\item Identity?
		\vskip 2em
		\item Commutative?
		\vskip 2em
	}
	Therefore, $(\mathbb{N},$\verb|^|$)$ is ...
\end{exercise}

\slide{
\begin{exercise}
	Let $*$ be the binary operation on $\mathbb{Z}$ given by \[a*b:=a-b.\]
	\enumalph{
		\item Associative?
		\vskip 2em
		\item Identity?
		\vskip 2em
		\item Commutative?
		\vskip 2em
	}
	Therefore, $(\mathbb{Z},-)$ is ...
\end{exercise}
}
\slide{
\begin{exercise}
	Let $*$ be the binary operation on $GL_2(\R)$given by \[A*B:=AB.\]
	\enumalph{
		\item Associative?
		\vskip 2em
		\item Identity?
		\vskip 2em
		\item Commutative?
		\vskip 2em
	}
	Therefore, $(GL_2(\R),\cdot)$ is ...
\end{exercise}
}
\slide{
	\begin{defn}
		Let $(M,*)$ be a monoid.

		If $x\in M$, we call $y\in M$ an \emph{inverse of $x$} if
			\[xy=e=yx.\]
		An element that has an inverse is called a \emph{unit}.
	\end{defn}
}
\slide{\vskip -.25in
	\begin{exercise}
		Fill in the table

			\def\arraystretch{1.5}
			$\begin{array}{c|c|c}
			\text{Monoid}&\quad e\quad &\quad \text{Set of Units}\quad\\
			\hline\hline
			(\Z,+)&&\\\hline
			(\R,+)&&\\\hline
			(\Z_n,+)&&\\\hline
			(M_n(\R),+)&&\\\hline
			(M_n(\R),\cdot)&&\\\hline
			(\R,\cdot)&&\\\hline
			(\Z_4,\cdot)&&\\\hline
			(\Z_n,\cdot)&&\\\hline
			(\Z_p,\cdot)&&\\\hline
			(S_3,\circ)&&\\\hline
			\end{array}$

		\begin{picture}(0,0)
		\put(195,132){$\leftarrow$ Matrices with real entries}
		\put(195,30){$\leftarrow$ $p$ is prime}
		\end{picture}
	\end{exercise}
}

\end{document}

